---
title: "Infected/Non-Infected Classification"
author: "Jacob Mardick"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **Libraries**

```{r Libraries, message=FALSE, warning=FALSE, echo=FALSE}
library(readxl)
library(sqldf)
library(dplyr)
library(tidyr)
```

## **User Input**

```{r User Input}
mz_of_interest <- c(721.5025, 722.5059, 722.513, 723.5164, 733.5025, 734.5059, 734.5342, 735.5375, 747.5182, 748.5216, 748.528, 749.532, 764.5236, 765.5269)
infected_dir <- c("C:/Users/Jacob/Box/Eberlin_Lab_BCM/People/MANOJ/INFECTED SAMPLE EXCEL SHEET/INFCTED SAMPLE 2023", "C:/Users/Jacob/Box/Eberlin_Lab_BCM/People/MANOJ/INFECTED SAMPLE EXCEL SHEET/INFECTED SAMPLE 2024")
uninfected_dir <- "C:/Users/Jacob/Box/Eberlin_Lab_BCM/People/MANOJ/Pedsarcoma Project/Excel sheet"
normalization_method <- "tic" ## tic, median, medianlog, maxpeak, or none
```

```{r include = FALSE}
## create directory for output files
#out_dir <- paste0(format(Sys.time(), "%Y-%m-%d_%H.%M"),"_species_outputs")
#if(!file.exists(file.path(out_dir))) {
  #dir.create(file.path(out_dir)) 
  #}
```

## **Peak match**

``` {r Peak List, message=FALSE, warning=FALSE}
# Create the m/z interest interval data frame
mz_interest_interval <- data.frame(
  target_mz = mz_of_interest,
  minus_005 = mz_of_interest - 0.005,
  plus_005 = mz_of_interest + 0.005
)
```

## **Read in infected and uninfected samples and combine to same data frame**

```{r Samples}
## Define function to process uninfected directory without classes
process_directory <- function(directory, infected_status) {
  file_list <- list.files(path = directory, pattern = "*.xlsx", full.names = TRUE)

  ## Exclude specific files
  file_list <- file_list[!grepl("Blank|List for Mass matchning", file_list)]
  file_list <- file_list[!grepl("~\\$", basename(file_list))]
  file_name_list <- tools::file_path_sans_ext(basename(file_list))

  data_list <- lapply(seq_along(file_list), function(i) {
    df <- read_excel(file_list[i], col_names = TRUE, skip = 7)
    df$Infected <- infected_status
    return(df)
  })

  ## Set names of the samples
  named_data_list <- setNames(data_list, file_name_list)

  return(named_data_list)
}

## Process infected and uninfected directories
infected_data_list <- process_directory(infected_dir, infected_status = 1)
uninfected_data_list <- process_directory(uninfected_dir, infected_status = 0)

## Combine into one list
data_list <- c(infected_data_list, uninfected_data_list)
data_list <- lapply(data_list, as.data.frame)

## Check if required columns exist in each data frame
required_columns <- c("Mass", "Intensity", "Infected")

## Search for which samples were not properly read
for (i in seq_along(data_list)) {
  sample <- data_list[[i]]
  if (!all(required_columns %in% names(sample))) {
    stop(paste("Data frame", names(data_list)[i], "is missing required columns"))
  }
}
```

## **Match sample peaks to target peak intervals**

```{r Peak Matching}
## Initiate empty list
peak_matched_list <- list()

## Peak match from mz_interest_interval
for (i in seq_along(data_list)) {
  sample <- data_list[[i]]
  ## Pass mz_interest_interval explicitly to sqldf
  matched_mz <- sqldf('SELECT mz_interest_interval.target_mz, sample.Intensity, sample.Infected
                      FROM sample, mz_interest_interval
                      WHERE sample.Mass between mz_interest_interval.minus_005 AND mz_interest_interval.plus_005',
                      list(sample = sample, mz_interest_interval = mz_interest_interval))
  peak_matched_list[[i]] <- matched_mz
}

## Add file names to peak matched list
names(peak_matched_list) <- names(data_list)
```

## **Combine all samples into one data frame and sum intensities of duplicate peaks**

```{r Combine}
## Combine list of sample dfs into one large df
all_samples_df <- bind_rows(peak_matched_list, .id = "Sample")

## Check for duplicates
duplicate_check <- all_samples_df %>%
  group_by(Sample, target_mz) %>%
  filter(n() > 1)
print(duplicate_check)

## If there are duplicates, aggregate them
if (nrow(duplicate_check) > 0) {
  all_samples_df <- all_samples_df %>%
    group_by(Sample, target_mz) %>%
    summarize(Intensity = sum(Intensity), .groups = 'drop')
}

## Add infection status back to all_samples_df
infection_status <- data_list %>%
  bind_rows(.id = "Sample") %>%
  select(Sample, Infected) %>%
  distinct()

all_samples_df <- all_samples_df %>%
  left_join(infection_status, by = "Sample")

## Check for duplicates (again)
duplicate_check2 <- all_samples_df %>%
  group_by(Sample, target_mz) %>%
  filter(n() > 1)
print(duplicate_check2)

## Reshape all_samples_df to m/z values as columns
wide_all_samples_df <- all_samples_df %>%
  pivot_wider(names_from = target_mz, values_from = Intensity)

## Vector for infection status
yall <- wide_all_samples_df$Infected

## Remove infection column from wide_all_samples_df
wide_all_samples_df <- wide_all_samples_df[, -2]
```

## **Function to calculate prediction of infected/uninfected**

```{r infection prediction function}
predict_infection <- function(data, mz_pair, threshold) {
  ## Denote each m/z intensity
  mz1 <- as.character(mz_pair[1])
  mz2 <- as.character(mz_pair[2])
  
  ## Check if both m/z values exist in the data
  if (!(mz1 %in% names(data) & mz2 %in% names(data))) {
    return(rep(NA, nrow(data)))
  }
  
  ## Calculate the ratio
  ratio <- data[[mz2]]/data[[mz1]]
  
  ## Apply threshold for prediction
  prediction <- ifelse(ratio > threshold, 1, 0)
  return(prediction)
}
```

## **Define m/z pairs and ratio thresholds**

```{r mz pairs}
## m/z pairs for ratio test
mz_pairs <- list(
  c(721.5025, 722.5059),
  c(722.513, 723.5164),
  c(733.5025, 734.5059),
  c(734.5342, 735.5375),
  c(747.5182, 748.5216),
  c(748.528, 749.532),
  c(764.5236, 765.5269)
)

thresholds <- c(0.473, 0.508, 0.484, 0.487, 0.495, 0.530, 0.530)
```

## Predict infected/uninfected in dataset

```{r predict}
## Initialize data frame to store predictions
predictions_df <- data.frame(matrix(nrow = nrow(wide_all_samples_df), ncol = 0))

## Loop through each sample and the 7 m/z pairs
for (i in seq_along(mz_pairs)) {
  mz_pair <- mz_pairs[[i]]
  threshold <- thresholds[i]
  col_name <- paste0("Prediction_", mz_pair[1], "_", mz_pair[2])
  predictions_df[[col_name]] <- predict_infection(wide_all_samples_df, mz_pair, threshold)
}

## Use max function across row to make final prediction
predictions_df$Final_Prediction <- do.call(pmax, c(predictions_df, list(na.rm = TRUE)))

## Add predictions to original data frame
final_predictions_df <- wide_all_samples_df %>%
  bind_cols(predictions_df %>%
              select(Final_Prediction))

## Add actual infection status to data frame
final_predictions_df <- final_predictions_df %>%
  left_join(infection_status %>%
              select(Sample, Infected), by = "Sample")

print(final_predictions_df)
#write.csv(final_predictions_df, "InfectedvsNoninfected_Manoj_predictions.csv")
#write.csv(predictions_df, "InfectedvsNoninfected_Manoj_predictionsratios.csv")

## Remove rows where final prediction is NA
final_predictions_df <- final_predictions_df %>%
  filter(!is.na(Final_Prediction))
```

## Confusion Matrix

```{r confusion matrix}
## Create confusion matrix
confusion_matrix <- table(Actual = factor(final_predictions_df$Infected, levels = c(0, 1), 
                                          labels = c("Uninfected", "Infected")), 
                          Predicted = factor(final_predictions_df$Final_Prediction, levels = c(0, 1), 
                                             labels = c("Uninfected", "Infected")))
print(confusion_matrix)
```





















