---
title: "Ratio Lasso - Species Prediction"
subtitle: "Renamed Sample Files"
author: "Jacob Mardick"
date: "`r format(Sys.time(), '%B %d, %Y %H:%M')`"
knit: (function(inputFile, encoding) { 
      proj_name <- tools::file_path_sans_ext(basename(inputFile));
      out_dir <- file.path("outputs", paste0(proj_name, "_", Sys.Date()));
      if(!file.exists(out_dir)) {   dir.create(out_dir) };
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file=file.path(dirname(inputFile), 
                        out_dir, 
                        paste0(format(Sys.time(), "%Y-%m-%d_%H.%M"),"_", proj_name, ".html"))) 
                        })

output: 
  html_document:
    keep_md: yes
    df_print: paged
    toc: false
geometry: margin=0.5in
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">
.main-container {
max-width: 1600px;
margin-left: auto;
margin-right: auto;
}
</style>
```

```{css, echo=FALSE}
h1, h2, h3, h4, h5 {
font-size: 20px;
}

h1, h2, h3, h4, h5, p {
text-align: center;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE, cache = FALSE)
```

```{r libraries, message = FALSE, warning = FALSE}
library(knitr)
library(kableExtra)
library(ggpubr)

library(tidyverse)
library(readxl)
library(sqldf)
library(doParallel)
library(stringr)
library(fcluster)
library(eberlinfuncs)

library(caret) ## automates supervised learning (predictive modeling)
library(glmnet) ## for training, cross validation, and testing model
library(pROC) ## for plotting ROC curve
library(scales) ## for integer y-axis on histogram
```

```{r user input}
## Full path to folder with sample files (excel or csv, or raw Thermo files)
sample_dir <- gsub("\\\\", "/", r"(C:\Users\Jacob\Box\Eberlin_Lab_BCM\People\MANOJ\Renamed_Gram_positive_Gram_negative_Isolates_May2025\Excel files)")

## Full path to feature list, otherwise NULL
feature_file <- gsub("\\\\", "/", r"(C:\Users\Jacob\Box\Eberlin_Lab_BCM\People\MANOJ\Renamed_Gram_positive_Gram_negative_Isolates_May2025\Manuscript_Microbes_Jan_2026\tentative_PNAS_Microbes_Supplementary_Data_14_Feb_2026.xlsx)")

## Full path to background peak list, otherwise NA
background_file <- NA

SNR_thresh <- NULL

ppm_error <- 5

mass_range <- c(100,1000)

## Peak Alignment Method: "clustering", "binning", or "featurelist"
peak_alignment_method <- "featurelist"

## If peak alignment method is "clustering": height at which to cut dendrogram to determine clusters
clust_h <- 0.01 

## Normalization Method: "tic", "maxpeak", "median", "medianlog", or "none"
normalization_method <- "tic" 

## ----FOR LASSO MODELS-----
train_fraction <- 0.7
seed <- 1234
```

```{r create directory for output files, include = FALSE}
proj_name <- tools::file_path_sans_ext(basename(rstudioapi::getSourceEditorContext()$path))

out_dir <- file.path("outputs", paste0(proj_name, "_", Sys.Date()))

if(!file.exists(out_dir)) {   
  dir.create(out_dir, recursive = TRUE) 
  }

files_dir <- file.path(out_dir, paste0(format(Sys.time(), "%Y-%m-%d_%H.%M"), "_", proj_name, "_files"))

if(!file.exists(files_dir)) {   
  dir.create(files_dir, recursive = TRUE) 
  }
```

```{r class_df}
class_df <- pull_classes(sample_dir)
class_df <- class_df %>%
  mutate(genus = sapply(str_split(class_df$V11, " "), '[[', 1)) %>%
  relocate(genus, .after = V10)

## class_df$sample_name <- class_df[cbind(1:nrow(class_df), max.col(!is.na(class_df), ties.method = 'last'))]

spec <- process_ms_files(sample_dir, ppm = ppm_error, mass_range = mass_range)
```

```{r peak alignment clustering}
## If peak alignment method is clustering
if (peak_alignment_method == "clustering") {
  
  ## Handling of multiple intensities aggregating to one cluster centroid: "sumints" or "maxint"
  clust_int_method  <-  "sumints" 
  
  ## Make vector of all sample m/z
  all_mz <- sort(unlist(
    lapply(spectra_list, function(y)
      lapply(y, function(x)
        round(x$mass, 3)))))
  
  names(all_mz) <- all_mz
  
  ## Cluster peaks into centroids
  tree <- fcluster(all_mz)
  clust_mz <- fcutree(all_mz, tree, h=clust_h)
  clustMZ <- sort(clust_mz$cen)
  
  ## Match cluster centroids to sample peaks
  clusterMatrixList <- lapply(spectra_list, function(x, y, z, w) 
    get_cluster_matrix(x, y, z, w), y=clustMZ, z = clust_h, w = clust_int_method)
  
  preprocList <- lapply(clusterMatrixList, function(x) as.matrix(get_data_matrix_clustering(x)))
  
  aligned_spectra <- do.call(rbind, preprocList)
  
  colnames(aligned_spectra) <- clustMZ
  rownames(aligned_spectra) <- sample_names
  
  ## FILTER #3: REMOVE RARE/UNCOMMON PEAKS PRESENT < 10% SAMPLES
  mz_count_filter <- colSums(aligned_spectra != 0) > as.integer(nrow(aligned_spectra)*0.10)
  aligned_spectra <- aligned_spectra[, mz_count_filter]
  
  filtered_mz <- clustMZ[mz_count_filter]
  
  ## Which sample peaks form the clusters
  clusters_df <- data.frame(cluster_id = clust_mz$cluster_index,
                            all_mz = all_mz)
  
  temp_a <- data.frame(clust_centroid = clust_mz$centroid, 
                       cluster_id = unique(clust_mz$cluster_index)[order(unique(clust_mz$cluster_index))])
  
  clusters_df <- merge(clusters_df, temp_a, by = "cluster_id")
  rm(temp_a)
  
  ## Remove duplicate rows of same cluster, mz, and centroid
  clusters_df <- clusters_df[!duplicated(clusters_df), ]
  
  ## only keep filtered_mz in min_max_clusters df
  clusters_df <- clusters_df[which(clusters_df$clust_centroid %in% filtered_mz), ]
  
  ## min and max for each centroid
  clusters_df <- clusters_df %>%
    group_by(cluster_id) %>%
    summarise(
      centroid = first(clust_centroid),
      min_mass = min(all_mz),
      max_mass = max(all_mz)
    )
  
  ## FILTER #4: REMOVE BACKGROUND PEAKS (IF BACKGROUND FILE PRESENT)
  if (!is.na(background_file)) {
    
    ## Check file extension
    if (file_ext(background_file) == "xlsx") {
      bg_mz <- data.frame(mass = unlist(read_excel(background_file)))
    } else if (file_ext(background_file) == "csv") {
      bg_mz <- data.frame(mass = unlist(read.csv(background_file, header = TRUE)))
    } else {
      print("Background file extension not recognized")
    }
    
    ## match bg_mz to cluster centroid
    bg_centroids <- sqldf("SELECT clusters_df.centroid, bg_mz.*
          FROM bg_mz,clusters_df
          WHERE bg_mz.mass BETWEEN clusters_df.min_mass AND clusters_df.max_mass")
    
    ## Remove peaks from aligned_spectra that are in bg_mz_centroid
    bg_ind <- which(filtered_mz %in% unique(bg_centroids$centroid))
    
    filtered_mz <- filtered_mz[-c(bg_ind)]
    aligned_spectra <- aligned_spectra[, -c(bg_ind)]
    
    ## only keep filtered_mz in min_max_clusters df
    clusters_df <- clusters_df[which(clusters_df$centroid %in% filtered_mz), ]
  }
}
```

```{r peak alignment binning}
if (peak_alignment_method == "binning") {
  features <- mode_binning(spectra_list = spec)
  feature_mean_or_mode <- "mean"
  aligned_spectra <- bin_spectra(spectra_list = spec, features, type = feature_mean_or_mode, class_df = class_df)
  filtered_mz <- switch(feature_mean_or_mode,
                        mean = features$mean_mz,
                        mode = features$mode_mz)
}
```

```{r peak alignment feature list}
if (peak_alignment_method == "featurelist") {
  
  feature_df <- readxl::read_xlsx(feature_file, sheet = "Sheet1", skip = 1)
  #features <- unlist(read_excel(feature_file, col_names = TRUE)[,1])
  features <- feature_df %>%
    filter(annotation_method == "MS2") %>%
    pull(mz) %>%
    as.numeric()
  
  ## subset features between mass range
  features <- features[features >= mass_range[1] & features <= mass_range[2]]
  
  matched_spec <-  target_mz_alignment(spectra_list = spec, target_mz = features, ppm = ppm_error, handle_dups = "sum")
  
  aligned_spectra <- bind_rows(matched_spec, .id = "sample_name") %>%
    pivot_wider(names_from = target_mz, values_from = intensity, values_fill = 0) %>%
    column_to_rownames(var = "sample_name")
  
  filtered_mz <- features
}
```

```{r normalization}
xall <- normalize_pixel(aligned_spectra, normalization_method)
```

## Create data for training

## Train/Test Split

``` {r split}
sample_names_df <- read.csv("Data/sample_names_df.csv")

class_df$patient <- sample_names_df$patient[match(class_df$V13, sample_names_df$sample_name)]

test_count <- round(table(class_df$V11)*(1-train_fraction),0)

test_index <- vector()
for (c in unique(class_df$V11)){
  set.seed(seed)
  test_index <- c(test_index, sample(which(class_df$V11 == c & is.na(class_df$patient)), 
                                     size = test_count[c]))
}

train_index <- (1:nrow(sample_names_df))[-test_index]

class_df$train_test_set <- ifelse(rownames(class_df) %in% train_index, "train", "test")

## test on isolates of matched specimens 1046, 1275, 1277, 1359, 1063
class_df$train_test_set[which(class_df$patient %in% c(1046, 1275, 1277, 1359, 1063))] <- "test"

otr <- which(class_df$train_test_set == "train")
sample_names_df$train_test_set <- ifelse(sample_names_df$X %in% otr, "train", "test")
yall <- factor(class_df$V11)
classes <- levels(yall)
```

## Below taken from Rob rc2 script for ratio calculations prior to Lasso

```{r all pairs}
## Function to compute log ratios with bounds
compute_logratios_with_bounds <- function(x, feature_names = NULL, min_valid = 10) {
  n_features <- ncol(x)
  n_samples <- nrow(x)
  n_ratios <- choose(n_features, 2)
  
  ## Initialize
  logratios <- matrix(NA_real_, nrow = n_samples, ncol = n_ratios)
  ratio_names <- character(n_ratios)
  
  ii <- 0
  for (j in 1:(n_features - 1)) {
    for (k in (j+1):n_features) {
      ii <- ii + 1
      ratio <- log(x[, j] / x[, k])
      
      ## Identify -Inf and Inf
      neg_inf_mask <- is.infinite(ratio) & ratio < 0
      pos_inf_mask <- is.infinite(ratio) & ratio > 0
      ratio[neg_inf_mask | pos_inf_mask] <- NA
      
      valid_vals <- ratio[!is.na(ratio)]
      
      ## Calculate quantiles per ratio
      if (length(valid_vals) >= min_valid) {
        q01 <- quantile(valid_vals, 0.01)
        q99 <- quantile(valid_vals, 0.99)
        
      } else {
        NA
      }
      
      ## Restore directionality of -Inf and +Inf
      ratio[neg_inf_mask] <- q01
      ratio[pos_inf_mask] <- q99
      ratio[is.na(ratio)] <- 0
      
      logratios[, ii] <- ratio
      if (!is.null(feature_names)) {
        ratio_names[ii] <- paste0(feature_names[j], "/", feature_names[k])
      } else {
        ratio_names[ii] <- paste0("V", j, "/V", k)
      }
    }
  }
  
  colnames(logratios) <- ratio_names
  return(list(logratios = logratios, ratio_names = ratio_names))
}

## Perform log ratio lasso
res <- compute_logratios_with_bounds(x = xall, feature_names = colnames(xall))
allpairs <- res$logratios
nams.p <- res$ratio_names

## Remove any features with remaining NAs
na_free_cols <- colSums(is.na(allpairs)) == 0
allpairs <- allpairs[, na_free_cols, drop = FALSE]
nams.p <- nams.p[na_free_cols]

## Keeping naming scheme the same
xtrain <- allpairs[otr,]
xtrain <- as(xtrain, "dgCMatrix")
xtest <- allpairs[-otr,]
ytrain <- yall[otr]
ytest <- yall[-otr]

model <- glmnet(
  xtrain, 
  ytrain, 
  family = "multinomial", ## for logit (logarithm of the odds) or logistic regression
  type.multinomial = "ungrouped",
  alpha = 1,
  standardize = FALSE, ## because intensities are already in same units
  lambda.min.ratio = 1e-05 ) ## The model will compute its own lambda sequence based on nlambda and lambda.min.ratio 

#nfolds <- nrow(xtrain)
nfolds <- 10
cvmodel <- cv.glmnet(xtrain, ytrain, nfolds = nfolds, family = "multinomial", type.measure = "class", standardize = F, keep = T, lambda.min.ratio = 1e-05, type.multinomial = "ungrouped")
```

```{r ROC Metrics}
## save index of lambda value that gives minimum cvm (mean cross-validated error)
min_lambda_index <- which(cvmodel$lambda == cvmodel$lambda.min)

## 1/(1+e^(-preval)) is the inverse of the link function?
cv_predictions <- 1/(1+exp(-cvmodel$fit.preval[, , min_lambda_index]))
```

```{r CV Confusion Matrix}
## predict classes based on threshold
cv_p_class <- apply(cv_predictions, 1, which.max)

cv_p_class <- factor(cv_p_class,levels=as.character(1:length(classes)),labels = classes)

cv_cm <- table(True=ytrain, Predict=cv_p_class)
```

```{r Training Histogram}
train_df <- data.frame(file_name=class_df$V13[otr], 
                       true_class=ytrain, 
                       prediction_class=cv_p_class, 
                       probability=cv_predictions)

rownames(train_df) <- NULL

train_wrong_preds <- train_df %>%
  filter(true_class != prediction_class)
```

```{r Report Model Coefficients}
lasso_coef <- reportCoef(allpairs, model, cvmodel$lambda.min, nams.p, classes)
```

```{r Testing}
test_p <- predict(model,
                  xtest,
                  s = cvmodel$lambda.min, ## Value of the penalty parameter lambda 
                  ## at which predictions are required
                  type = "response" ) ## to get prediction values rather than linker function values
```

```{r Testing Confusion Matrix}
test_p_class <- apply(test_p, 1, which.max)

test_p_class <- factor(test_p_class,levels=as.character(1:length(classes)),labels = classes)

test_cm <- table(True=ytest,
                 Predict=test_p_class)
```

```{r Test Histogram}
test_df <- data.frame(file_name=class_df$V13[-c(otr)], 
                      true_class=ytest, 
                      prediction_class=test_p_class, 
                      probability=as.data.frame(test_p))

rownames(test_df) <- NULL

test_wrong_preds <- test_df %>%
  filter(true_class != prediction_class)
```

```{r save model, include=FALSE}
save(model, cvmodel, filtered_mz, mass_range, ppm_error, normalization_method, seed, classes, lasso_coef, peak_alignment_method,
     file=file.path(files_dir, paste0(format(Sys.time(), "%Y-%m-%d_%H.%M"),"_",proj_name, "_model.RData")))
```

```{r}
save(list = ls(), file=file.path(files_dir, paste0(format(Sys.time(), "%Y-%m-%d_%H.%M"),"_",proj_name, "_ALL.RData")))
```

```{r save files, include = FALSE}
## training set files

write.csv(cv_cm, 
          file = file.path(files_dir, "train_cm.csv"))

write.csv(train_df, 
          row.names = FALSE,
          file.path(files_dir, "train_preds.csv"))

write.csv(train_wrong_preds, 
          row.names = FALSE,
          file.path(files_dir, "misclassified_train_preds.csv"))

write.csv(as.matrix(lasso_coef), 
          file.path(files_dir, "lasso_coefficients.csv"))

## test set files

write.csv(test_cm, 
          file = file.path(files_dir, "test_cm.csv"))

write.csv(test_df, 
          row.names = FALSE,
          file.path(files_dir, "test_preds.csv"))

write.csv(test_wrong_preds, 
          row.names = FALSE,
          file.path(files_dir, "misclassified_test_preds.csv"))
```

```{r save plots, include=FALSE}
png(filename=file.path(files_dir, "cv_plot.png"))
plot(cvmodel)
title("cross-validation curve", line = 2.5)
dev.off()
```

<br>

#### **Preprocessing and Statistical Model Settings**

```{r chunk3, fig.align = "center"}
if (peak_alignment_method == "clustering") {
  cluster_bin_size <- c("Cluster Height:", clust_h)
} else if (peak_alignment_method == "binning") {
  cluster_bin_size <- c("Bin Size:", "0.01")
} else if (peak_alignment_method == "featurelist") {
  cluster_bin_size <- c("Peak Mass Error:", paste0(ppm_error, " ppm"))
}

if (is.na(background_file)) {
  bg_exclusion <- "no"
} else if (!is.na(background_file)) {
  bg_exclusion <- "yes"
}

# if (is.na(SNR_thresh)) {
#   SNR_thresh <- "No noise information provided"
# } else if (!is.na(SNR_thresh)) {
#   SNR_threshold <- SNR_thresh
# }

settings_df <- rbind(#c("SNR Threshold:", SNR_thresh),
  c("Mass Range:", paste0('<i>m/z</i> ', mass_range[1], " - ", mass_range[2])),
  c("Peak Alignment Method:", peak_alignment_method),
  cluster_bin_size,
  c("Background Peak Exclusion:", bg_exclusion),
  c("Normalization Method:", normalization_method),
  c("Randomization Seed:", seed),
  c("Train/Test Split:", paste0((train_fraction*100),"/",(100-(train_fraction*100)))),
  c("Elastic Net Alpha:", "1 (LASSO)"))

species <- gsub("_", " ", classes)

kable(settings_df,
      row.names = FALSE,
      align = "l",
      format = "html",
      escape = FALSE)%>%
  column_spec(1:2, width = "3in")%>% 
  kable_styling(full_width = F)

cols <- rev(c(hue_pal()(2)))
```

```{r chunk4}
cv_recalls <- paste0(format(round(sapply(1:length(classes), function(x) cv_cm[x,x]/sum(cv_cm[x,])*100),2), nsmall = 2), "%")
cv_counts <- sapply(1:length(classes), function(x) paste0(cv_cm[x,x],"/",sum(cv_cm[x,])))
cv_accuracy <- cbind(paste0(species, ":"), cv_recalls, cv_counts)

cv_accuracy <- rbind(cv_accuracy, 
                     c("Overall Accuracy:", 
                       paste0(format(round(sum(diag(cv_cm))/sum(cv_cm)*100,2), nsmall = 2), "%"),
                       paste0(sum(diag(cv_cm)),"/",sum(cv_cm))))

colnames(cv_accuracy) <- c("Species", "Recall Rate", "Correct Prediction/True Count")


test_recalls <- paste0(format(round(sapply(1:length(classes), function(x) test_cm[x,x]/sum(test_cm[x,])*100),2), nsmall = 2), "%")
test_counts <- sapply(1:length(classes), function(x) paste0(test_cm[x,x],"/",sum(test_cm[x,])))

test_accuracy <- cbind(paste0(species, ":"), test_recalls,test_counts)

test_accuracy <- rbind(test_accuracy, 
                       c("Overall Accuracy:", 
                         paste0(format(round(sum(diag(test_cm))/sum(test_cm)*100,2), nsmall = 2), "%"),
                         paste0(sum(diag(test_cm)),"/",sum(test_cm))))
colnames(test_accuracy) <- c("Species", "Recall Rate", "Correct Prediction/True Count")
```

```{r chunk5}
coef2 <- format(round(lasso_coef,3), nsmall = 3)[ , 1:(ncol(lasso_coef)-2)]
coef2_feature_names <- rownames(coef2)
coef2_feature_names[1] <- "Intercept"
coef2 <- data.frame(Features = coef2_feature_names, coef2, row.names = NULL, check.names = FALSE)
```

<br>

#### **Training Set Confusion Matrix**

```{r fig.height=6, fig.width=6, message = FALSE, fig.align = "center"}
cv_cm_df <- as.data.frame(cv_cm)
cv_cm_plot_df <- cv_cm_df %>%
  mutate(plot.fill = case_when(
    True == Predict ~ "TP",
    True != Predict & Freq > 0 ~ "FP",
    True != Predict & Freq == 0 ~ "TN"))

ggplot(cv_cm_plot_df, aes(Predict, True, label = Freq, fill = plot.fill)) + 
  geom_tile(color = "black") +
  geom_text() +
  scale_fill_manual(values = c("#f2e3c7", "white", "#c3fcb5"))+
  scale_x_discrete(position = "top") +
  scale_y_discrete(limits = rev) +
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0), 
        plot.title.position = "plot",
        legend.position = "none",
        panel.grid = element_blank()) +
  #labs(title = "Training Set") +
  scale_y_discrete(labels=rev(species), limits=rev) +
  scale_x_discrete(labels=species, position = "top")
```

```{r, fig.align = "center"}
kable(cv_accuracy,
      row.names = FALSE,
      align = "r") %>% 
  column_spec(1, bold = TRUE) %>%
  kable_minimal(full_width = FALSE,
                bootstrap_options = "condensed",
                html_font = "Calibri",
                font_size = 16)
```

<br>

#### **Test Set Confusion Matrix**
```{r fig.height=6, fig.width=6, message = FALSE, fig.align = "center"}
test_cm_df <- as.data.frame(test_cm)
test_cm_plot_df <- test_cm_df %>%
  mutate(plot.fill = case_when(
    True == Predict ~ "TP",
    True != Predict & Freq > 0 ~ "FP",
    True != Predict & Freq == 0 ~ "TN"))

ggplot(test_cm_plot_df, aes(Predict, True, label = Freq, fill = plot.fill)) + 
  geom_tile(color = "black") +
  geom_text() +
  scale_fill_manual(values = c("#f2e3c7", "white", "#c3fcb5"))+
  scale_x_discrete(position = "top") +
  scale_y_discrete(limits = rev) +
  theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0), 
        plot.title.position = "plot",
        legend.position = "none",
        panel.grid = element_blank()) +
  #labs(title = "Test Set") +
  scale_y_discrete(labels=rev(species), limits=rev) +
  scale_x_discrete(labels=species, position = "top")
```

```{r, fig.align = "center"}
kable(test_accuracy,
      row.names = FALSE,
      align = "r") %>% 
  column_spec(1, bold = TRUE) %>%
  kable_minimal(full_width = FALSE,
                bootstrap_options = "condensed",
                html_font = "Calibri",
                font_size = 16)
```

<br>

#### **Model Predictive Feature Weights**
Positive value  indicates that the feature is predictive for that species, negative value indicates that the feature is predictive for the opposite class (not that species).

```{r, fig.align = "center"}
coef2 %>%
  kable(
    #caption = "LASSO Model Predictive Feature Weights",
    row.names = FALSE,
    align = "c") %>% 
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "condensed"), font_size = 12) %>%
  column_spec(1:ncol(coef2),width = "2in")
```

<br>

```{r chunk10}
## Add misclassified training samples
train_wrong_preds$true_class <- gsub("_", " ", train_wrong_preds$true_class)
train_wrong_preds$prediction_class <- gsub("_", " ", train_wrong_preds$prediction_class)
train_wrong_preds[, 4:ncol(train_wrong_preds)] <- round(train_wrong_preds[, 4:ncol(train_wrong_preds)], 3)

## Add misclassified test samples
test_wrong_preds$true_class <- gsub("_", " ", test_wrong_preds$true_class)
test_wrong_preds$prediction_class <- gsub("_", " ", test_wrong_preds$prediction_class)
test_wrong_preds[, 4:ncol(test_wrong_preds)] <- round(test_wrong_preds[, 4:ncol(test_wrong_preds)], 3)
```

<br>

#### **Training Set Misclassifications**

```{r chunk11, fig.align = "center"}
align <- paste0("lll",paste0(rep("c", ncol(train_wrong_preds)-3), collapse =""))

kable(train_wrong_preds,
      #caption = "Training Set Misclassifications",
      col.names = c("Sample Name", "True Class", "Predicted Class", species),
      row.names = FALSE,
      align = align) %>%
  kable_styling(full_width = FALSE, 
                position = "center", 
                bootstrap_options = c("striped", "condensed"), 
                font_size = 12) %>%
  column_spec(1:3, width_min = "2in", width_max = "5in")  %>%
  add_header_above(c(" " = 3, "Probability" = ncol(train_wrong_preds)-3))
```

<br>

#### **Test Set Misclassifications**

```{r chunk12, fig.align = "center"}
kable(test_wrong_preds,
      col.names = c("Sample Name", "True Class", "Predicted Class", species),
      row.names = FALSE,
      align = align)  %>%
  kable_styling(full_width = FALSE, 
                position = "center", 
                bootstrap_options = c("striped", "condensed"), 
                font_size = 12) %>%
  column_spec(1:3, width_min = "2in", width_max = "5in")  %>%
  add_header_above(c(" " = 3, "Probability" = ncol(test_wrong_preds)-3))
```

<br>

#### **Train/Test Data Split**

```{r chunk13}
## Add list of samples for training
## Add list of samples for testing

train_samples <- sort(train_df$file_name)

test_samples <- sort(test_df$file_name)

n <- max(length(train_samples), length(test_samples))
length(train_samples) <- n                      
length(test_samples) <- n

samples <- cbind("Training Samples" = train_samples,
                 "Test Samples" = test_samples)

samples[is.na(samples)] <- ""
```

```{r chunk14}
kable(samples,
      #caption = "Train/Test Data Split",
      row.names = FALSE,
      align = "l") %>%
  kable_minimal(full_width = FALSE,
                bootstrap_options = "condensed",
                html_font = "Calibri",
                font_size = 12)
```